---
title: "Conclusions"
editor: visual
format:
  html:
    toc: false
execute:
  echo: false
---

This project sought to answer a number of questions, but at their core, all pointed towards one particular curiosity --- Can Magic: The Gathering Cards Predict Hasbro's Stock Price?

In pursuit of an answer, this project used exploratory data analysis to establish the connection between Magic cards and Hasbro's stock, built several time series models for evaluating and forecasting Hasbro's stock prices, and considered deep learning techniques for building upon these initial efforts.

Ultimately, this analysis is too narrow in scope to answer this question sufficiently. One result that is apparent, however, is that if Magic: the Gathering Cards can predict Hasbro's stock price, it is through a more complicated model or set of mechanisms than are presented here in this small scope.

Along the way to this conclusion, several data constraints or other obstacles introduced limitations to the results. These challenges are worth further exploring and present a set of meaningful lessons and in and of themselves.

```{r}
# [Insert custom graphic of Hasbro stock TS and magic cards or something similar]
```

### Data Constraints

```{r}
# weekly data for magic the gathering
# magic card data had to be manually collected
# stocks are influenced by countless macro-economic and micro-economic features, even if magic is a large portion of Hasbro's portfolio, it is a poor proxy variable without a number of other features
# the volume of sales for magic products, player sentiment, and the supply of cards brought to market would all be helpful variables, but unavailable for one reason or another
```

### Findings

```{r}
# Theory of multi-variate analysis improving on ARMA model.
# Magic cards seemed to play a role in that improvement.
# Granularity of data is crucial.
# Pros and Cons of Sealed products as Magic-stand in

```
